Sampling Strategies in Decisions from Experience

### 2021-11-03 manuscript.Rmd
Review of data visualization

```{r}
# load packages
pacman::p_load(tidyverse, 
               knitr, 
               scico # for scientific color palettes
               )
```

This document was created from the commit with the hash ``r repro::current_hash()``

# Author Note

All materials of this work in progress can be found on the accompanying GitHub repository: https://github.com/linushof/sampling-in-dfe

- Add information on how to reproduce the project

- Add contact

# Abstract

A probability theoretic definition of prospects and a rough stochastic sampling model for decisions from experience is proposed.
It is demonstrated how the model can be used a) to explicate assumptions about the sampling and decision strategies that agents may apply and b) to derive predictions about function forms and parameter values that describe the resulting decision behavior. 
Synthetic choice data is simulated and modeled in cumulative prospect theory to test these predictions.

@@

## Sampling in Decisions from Experience

In research on the decision theory, a standard paradigm is the choice between at least two (monetary) prospects.
Let a prospect be a probability space $(\Omega, \mathscr{F}, P)$.
$\Omega$ is the sample space

$$\begin{equation}
\Omega = \{\omega_1, ..., \omega_n\}
\end{equation}$$

containing a finite set of possible outcomes $\omega$, monetary gains and/or losses respectively. 
$\mathscr{F}$ is the set of all possible subsets of $\Omega$:

$$\begin{equation}
\mathscr{F} = \{A_1, A_2, ...\} = \mathscr{P}(\Omega) 
\; .
\end{equation}$$

@@

$$\begin{equation}
P: \mathscr{F} \mapsto [0,1] 
\end{equation}$$

@@

(1) $X$ is a measurable function $$\begin{equation} X: (\Omega, \mathscr{F})  \mapsto (\Omega', \mathscr{F'}) \; , \end{equation}$$ where $\Omega'$ is a set of real numbered values $X$ can take and $\mathscr{F'}$ is a set of subsets of $\Omega'$. I.e., $\Omega$ maps into $\Omega'$ such that correspondingly each subset $A' \in \mathscr{F'}$ has a pre-image $X^{-1}A' \in \mathscr{F}$, which is the set $\{\omega \in \Omega: X(\omega) \in A'\}$ [@kolmogorovFoundationsTheoryProbability1950, p. 21].

@@

$$\begin{equation}
D:= f((\Omega_1, \mathscr{F}_1, P_1), ..., (\Omega_k, \mathscr{F}_k, P_k))
\;.
\end{equation}$$

@@

$$\begin{equation}
D := f(X_{i1}, ..., X_{ik}) 
\; ,
\end{equation}$$

@@

## Capturing Differences in Sampling and Decision Strategies

@@

Let $X$ and $Y$ be random variables defined on the prospects $(\Omega_X, \mathscr{F}_X, P_X)$ and $(\Omega_Y, \mathscr{F}_Y, P_Y)$.

@@

$$\begin{equation}
\mathscr{C} = \left\{B_1 = \{c \in C: \overline{X}_{N_X} - \overline{Y}_{N_Y} > 0\}, B_2 = \{ c \in C: \overline{X}_{N_X} - \overline{Y}_{N_Y} \leq 0\}
\right\}
\end{equation}$$

be a set of subsets of $C$, indicating that comparisons of prospects on the ordinal (rather than on the metric) scale are of primary interest. 
We denote a single ordinal comparison by the event space $(C, \mathscr{C})$.
The observed outcome of such an ordinal comparison can be regarded as evidence for or against a prospect and the number of wins over a series of independent ordinal comparisons as accumulated evidence. 
To integrate the concept of evidence accumulation into the current model, we let the decision variable $D$ be a measurable function

$$\begin{equation}
D: C \mapsto C'
\end{equation}$$

that maps the possible outcomes of a mean comparison in $C$ onto a measure space $C' = \{0,1\}$, with $0$ ($1$) indicating a lost (won) comparison:

$$\begin{equation}
D(c \in C) \in C' = 
  \begin{cases}
    1 & \text{for} & \{c \in C: (\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0)\}  \in \mathscr{C} \\
    0 & \text{else}. 
  \end{cases} 
\end{equation}$$

@@

$$\begin{equation}
D \sim \mathcal{B}\left( p \left(\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0\right), n\right) \; ,
\end{equation}$$

@@

I.e., first, whether a single (summary decision) or multiple ordinal comparisons (round-wise decision) are conducted, is reflected by the length of the sequence $D_i = D_1, ..., D_n$, with $n = 1$ for summary decisions and $n \geq 1$ for round-wise decisions. 
The idea that multiple comparisons may be conducted can be formalized by extending the event space $(C, \mathscr{C})$ of a single comparison to its $n$-fold product

$$\begin{equation}
\left(\prod_{i = 1}^{n} C_i, \bigotimes_{i = 1}^{n} \mathscr{C}_i \right)
\; ,
\end{equation}$$

@@

$$\begin{equation}
\left(\prod_{i = 1}^{n} C_i, \bigotimes_{i = 1}^{n} \mathscr{C}_i, \bigotimes_{i = 1}^{n} P_i \right)
\; ,
\end{equation}$$

including the probability measure $P$, which assigns each $c \in C$, and each $B \in \mathscr{C}$ respectively, a probability, must be identified. 
Although this can be demonstrated for the random variables considered so far, it becomes rather quickly intractable with increasing elements in $\Omega$ and growing sequence lengths. 
However, by explicating the above probability measure ([Appendix 1][Appendix 1: Probability Measure Over Sample Mean Comparisons]), it becomes evident that it is largely influenced by the underlying sequence lengths $N_X$ and $N_Y$.
In the following section and a subsequent simulation analysis, we demonstrate that this influence is of such a form that it produces some principle and predictable trends in the choice behavior that can be modeled in the cumulative prospect theory.

## Predicting Choice Behavior

@@

For each distinct strategy-parameter combination, we ran 20 chains of 20,000 iterations each, after a warm-up period of 1000 samples.
To reduce potential autocorrelation during the sampling process, we only kept every 20th sample (thinning). 
Bar plots displaying the potential scale reduction factors $\hat{R}$ [@gelmanInferenceIterativeSimulation1992] of the group and agent level CPT parameters can be found in [Appendix 3][Appendix 3: Convergence Diagnostics for CPT Parameters].

@@

trial_length %>%
  ggplot(aes(x = s, y = med)) +
  geom_jitter(color = "#CECECE", size = 3) +
  geom_point(data = trial_length_s, aes(color = s), size = 3) +
  geom_path(data = trial_length_s, aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 120, 10)) +
  labs(x ="Switching Probability",
       y = "Median Trial Length", 
       color="Switching Probability") + 
  theme_minimal()
```

@@

# call script for preparing CPT data

source("R/prepare_estimates.R") 
gel_92 <- cpt %>% select(s, boundary, a, parameter, level, Rhat) # store scale reduction factor

# Tidy: CPT parameters in columns

cpt <- cpt %>% 
  select(s, boundary, a, parameter, level, mean) %>% 
  pivot_wider(names_from = parameter, values_from = mean)

# agent level estimates

cpt_agent <- cpt %>% filter(level != "mu") # agent level estimates

## compute mean across all agent level estimates

cpt_agent_mean <- cpt_agent %>% 
  group_by(s, boundary, a) %>% 
  summarise(alpha = round(mean(alpha), 2), 
            gamma = round(mean(gamma), 2), 
            delta = round(mean(delta), 2), 
            rho = round(mean(rho), 2))

# group level estimates

cpt_mu <- cpt %>% filter(level == "mu") # group level posterior means
```

@@

# gamma

cpt_agent %>% 
  ggplot(aes(x = s, y = gamma)) +
  geom_jitter(color = "#CECECE" ) + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 3) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Curvature ", gamma)),
       x ="Switching Probability",
       y = expression(gamma), 
       color="Switching Probability") + 
  theme_minimal()

# delta

cpt_agent %>% 
  ggplot(aes(x = s, y = delta)) +
  geom_jitter(color = "#CECECE") + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 3) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 10, .5)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Elevation ", delta)),
       x ="Switching Probability",
       y = expression(delta), 
       color="Switching Probability") + 
  theme_minimal()

# weighting function

@@

cpt_agent_w <- cpt_agent %>% 
  select(-c(alpha, rho)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>% # 
  mutate(w = round(exp(-delta*(-log(ep))^gamma), 2))

cpt_agent_mean_w <- cpt_agent %>% 
  group_by(s) %>% 
  summarise(gamma = round(mean(gamma), 2), 
            delta = round(mean(delta), 2)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>%
  mutate(w = round(exp(-delta*(-log(ep))^gamma), 2))

@@

cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w)) +
  geom_path(color = "#CECECE") + 
  geom_path(data = cpt_agent_mean_w, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()

## for separate a

cpt_agent_mean_w <- cpt_agent %>% 
  group_by(s, a) %>% 
  summarise(gamma = round(mean(gamma), 2), 
            delta = round(mean(delta), 2)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>%
  mutate(w = round(exp(-delta*(-log(ep))^gamma), 2))

cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w)) +
  geom_path(color = "#CECECE") + 
  geom_path(data = cpt_agent_mean_w, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  facet_wrap(~a, nrow = 2) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()
```

@@

@@

@@

We start by noting that the probability of success equals the probability of the subset $B_1 \in \mathscr{C}$ which contains all comparisons $c \in C$ that satisfy the inequation $\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0$.
Since the probability of the subset $B_1$ equals the probability that any of its elements occurs, the probabilities of all $c \in B_1$ are summed up. 
We therefore use the sum operator with the limits $j \in \{1, ..., |B_1|\}$, with $|B_1|$ denoting the cardinality of the set $B_1$, i.e., the number of elements within the set. 
As a major assumption is that the random variables within a sample sequence are iid and independent across sequences, the joint probability of a pair $\overline{X}_{N_X} \cap \overline{Y}_{N_Y}$, i.e., a pair $X_i \cap Y_i$ with a unique combination of the number of times any possible realization of $X$ and $Y$ occurs, is described by the product of their individual probabilities $p(\overline{X}_{N_X})$ and $p(\overline{Y}_{N_Y})$.
Finally, to get these individual probabilities, we must determine the probability that each possible realization indeed occurs the respective number of times in the underlying sample sequence.
This probability is given by the multinomial distribution which reduces to a binomial distribution in the case of only two possible realizations. 
We denote the number of times a realization occurs in a sample sequence with $k_{\omega}$, where $\omega$ signals that the realizations of the random variables are numerical equivalent to the outcomes in the sample space of the prospects.

$$\begin{equation}
  \begin{split}
    p(D = 1) 
      & = p \left( \; B_1 = \{c \in C: (\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0)\} \in \mathscr{C} \; \right) \\\\\\ 
      & = \sum_{j=1}^{|B_1|} \left( \;
          p(\overline{X}_{N_Xj})  
          \times 
          p(\overline{Y}_{N_Yj}) \; 
        \right)_{j \, \in \, \{1, ..., |B_1|\}} \\\\\\
      & = \sum_{j=1}^{|B_1|} \left( \; 
        p \left( 
          \frac{1}{N_X} \sum\limits_{i = 1}^{N_X} X(\omega \in \Omega_X)_{ij}
        \right) 
        \times
        p \left( 
          \frac{1}{N_Y} \sum\limits_{i = 1}^{N_Y} Y(\omega \in \Omega_Y)_{ij}
        \right) \; 
        \right)_{j \, \in \, \{1, ..., |B_1|\}} \\\\\\
      & = \sum_{j=1}^{|B_1|} \left( \;
        \left(
          {N_X\choose k_{\omega_{1X}j}, ..., k_{\omega_{nX}j}} \prod_{\omega_{iX} \in \Omega_X} p(\omega_{iX})^{k_{\omega_{iX}j}}
        \right)
        \times
        \left(
        {N_Y\choose k_{\omega_{1Y}j}, ..., k_{\omega_{nY}j} } \prod_{\omega_{iY} \in \Omega_Y} p(\omega_{iY})^{k_{\omega_{iY}j}}
        \right)
        \right)_{j \, \in \, \{1, ..., |B_1|\}} \\\\\\
      & = \sum_{j=1}^{|B_1|} \left( \;
        \left(
          \frac{N_X!}{k_{\omega_{1X}j}! \times ... \times k_{\omega_{nX}j}!} 
          \times 
          p(\omega_{1X})^{k_{\omega_{1X}j}} 
          \times 
          ... 
          \times p(\omega_{nX})^{k_{\omega_{nX}j}}
        \right)
        \times
        \left(
          \frac{N_Y!}{k_{\omega_{1Y}j}! \times ... \times k_{\omega_{nY}j}!} 
          \times 
          p(\omega_{1Y})^{k_{\omega_{1Y}j}} 
          \times 
          ... 
          \times p(\omega_{nY})^{k_{\omega_{nY}j}}
        \right)
      \right)_{j \, \in \, \{1, ..., |B_1|\}}
  \end{split}
\end{equation}$$

@@

## Appendix 3: Convergence Diagnostics for CPT Parameters

```{r}
# group level

gel_92 %>% 
  filter(level == "mu") %>% 
  ggplot(aes(x = boundary, y = Rhat, fill = s)) +
  geom_bar(position = "dodge_2", stat = "identity") +
  scale_fill_scico(palette = "buda") + 
  scale_x_discrete(name = element_blank()) +
  scale_y_continuous(breaks = seq(.1, 1.5, .2)) +
  geom_hline(yintercept = 1, color = "black") +
  geom_hline(yintercept = 1.1, color = "red", linetype = "dashed") +
  facet_grid(parameter~a, switch = "y") +
  labs(y = "Scale Reduction Factor", 
       fill="Switching Probability") + 
  theme_minimal()

# agent level

parameters <- c("alpha", "gamma", "delta", "rho")
for(i in seq_along(parameters)) {

print(gel_92 %>% 
  filter(level != "mu", parameter == parameters[i]) %>% 
  ggplot(aes(x = level, y = Rhat, fill = s)) +
  geom_bar(stat = "identity", position = "dodge_2") +
  scale_x_discrete(breaks = element_blank(), name = element_blank()) + 
  scale_y_continuous(breaks = seq(.1, 1.5, .2)) +
  geom_hline(yintercept = 1, color = "black") +
  geom_hline(yintercept = 1.1, color = "red", linetype = "dashed") +
  facet_grid(boundary~a, switch = "y") +
  labs(title = parameters[i],
       y = "Scale Reduction Factor", 
       fill="Switching Probability") + 
  scale_fill_scico(palette = "buda") + 
  theme_minimal())
}
```

>>>>

```{r}
# load packages
pacman::p_load(tidyverse, 
               knitr, 
               ggbeeswarm,
               scico # for scientific color palettes
               )
```

@@

# Sampling in Decisions from Experience

In research on the decision theory, a standard paradigm is the choice between at least two (monetary) prospects.
Let a prospect be a probability space $(\Omega, \mathcal{F}, P)$.
$\Omega$ is the sample space

$$
\Omega = \{\omega_1, ..., \omega_n\}
$$

containing a finite set of possible outcomes $\omega$, monetary gains and/or losses respectively. 
$\mathcal{F}$ is the set of all possible subsets of $\Omega$:

$$
\mathcal{F} = \{A_1, A_2, ...\} = \mathcal{P}(\Omega) 
\; .
$$

@@

$$
P: \mathcal{F} \mapsto [0,1] 
$$

@@

(1) $X$ is a measurable function $$X: (\Omega, \mathcal{F})  \mapsto (\Omega', \mathcal{F'}) \; ,$$ where $\Omega'$ is a set of real numbered values $X$ can take and $\mathcal{F'}$ is a set of subsets of $\Omega'$. I.e., $\Omega$ maps into $\Omega'$ such that correspondingly each subset $A' \in \mathcal{F'}$ has a pre-image $X^{-1}A' \in \mathcal{F}$, which is the set $\{\omega \in \Omega: X(\omega) \in A'\}$ [@kolmogorovFoundationsTheoryProbability1950, p. 21].

@@

$$
D:= f((\Omega_1, \mathcal{F}_1, P_1), ..., (\Omega_k, \mathcal{F}_k, P_k))
\;.
$$

@@

$$
D := f(X_{i1}, ..., X_{ik}) 
\; ,
$$

@@

# Capturing Differences in Sampling and Decision Strategies

@@

Let $X$ and $Y$ be random variables defined on the prospects $(\Omega_X, \mathcal{F}_X, P_X)$ and $(\Omega_Y, \mathcal{F}_Y, P_Y)$.

@@

$$
\mathcal{G} = \left\{B_1 = \{c \in C: \overline{X}_{N_X} - \overline{Y}_{N_Y} > 0\}, B_2 = \{ c \in C: \overline{X}_{N_X} - \overline{Y}_{N_Y} \leq 0\}
\right\}
$$

be a set of subsets of $C$, indicating that comparisons of prospects on the ordinal (rather than on the metric) scale are of primary interest. 
We denote a single ordinal comparison by the event space $(C, \mathcal{G})$.
The observed outcome of such an ordinal comparison can be regarded as evidence for or against a prospect and the number of wins over a series of independent ordinal comparisons as accumulated evidence. 
To integrate the concept of evidence accumulation into the current model, we let the decision variable $D$ be a measurable function

$$
D: C \mapsto C'
$$

that maps the possible outcomes of a mean comparison in $C$ onto a measure space $C' = \{0,1\}$, with $0$ $(1)$ indicating a lost (won) comparison:

$$
D(c \in C) \in C' = 
  \begin{cases}
    1 \quad \text{for} \quad \{c \in C: (\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0)\}  \in \mathcal{G} \\
    0 \quad \text{else}. 
  \end{cases} 
$$

@@

$$
D \sim \mathcal{B}\left( p \left(\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0\right), n\right) \; ,
$$

@@

I.e., first, whether a single (summary decision) or multiple ordinal comparisons (round-wise decision) are conducted, is reflected by the length of the sequence $D_i = D_1, ..., D_n$, with $n = 1$ for summary decisions and $n \geq 1$ for round-wise decisions. 
The idea that multiple comparisons may be conducted can be formalized by extending the event space $(C, \mathcal{G})$ of a single comparison to its $n$-fold product

$$
\left(\prod_{i = 1}^{n} C_i, \bigotimes_{i = 1}^{n} \mathcal{G}_i \right)
\; ,
$$

@@

$$
\left(\prod_{i = 1}^{n} C_i, \bigotimes_{i = 1}^{n} \mathcal{G}_i, \bigotimes_{i = 1}^{n} P_i \right)
\; ,
$$

including the probability measure $P$, which assigns each $c \in C$, and each $B \in \mathcal{G}$ respectively, a probability, must be identified. 
Although this can be demonstrated for the random variables considered so far, it becomes rather quickly intractable with increasing elements in $\Omega$ and growing sequence lengths. 
However, by explicating the above probability measure ([Appendix 1][Appendix 1: Probability Measure Over Sample Mean Comparisons]), it becomes evident that it is largely influenced by the underlying sequence lengths $N_X$ and $N_Y$.
In the following section and a subsequent simulation analysis, we demonstrate that this influence is of such a form that it produces some principle and predictable trends in the choice behavior that can be modeled in the cumulative prospect theory.

# Predicting Choice Behavior

@@

# read cpt data

cols <- list(.default = col_double(),
             boundary = col_factor(),
             a = col_factor(),
             parameter = col_factor())
cpt_long <- read_csv("data/estimates/estimates_cpt.csv", col_types = cols)

# store convergence diagnostics

gel_92 <- cpt_long %>% select(s, boundary, a, parameter, Rhat, n.eff) 
```

For each distinct parameter combination, we ran 20 chains of 40,000 iterations each, after a warm-up period of 1000 samples.
To reduce potential autocorrelation during the sampling process, we only kept every 20th sample (thinning).
The potential scale reduction factor $\hat{R}$ [@gelmanInferenceIterativeSimulation1992] was $\leq$ `r round(max(gel_92$Rhat), 3)` for all parameters, indicating good convergence.
The minimum effective sample size was `r min(gel_92$n.eff)`.

@@

trial_length %>%
  ggplot(aes(x = s, y = med)) +
  geom_jitter(color = "#CECECE", size = 3) +
  geom_point(data = trial_length_s, aes(color = s), size = 3) +
  geom_path(data = trial_length_s, aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 120, 10)) +
  labs(title = "Trial Length", 
       x ="Switching Probability",
       y = "Median Trial Length", 
       color="Switching Probability") + 
  theme_minimal()
```

@@

# tidy CPT data: parameters as separate columns

cpt_wide <- cpt_long %>% 
  select(s, boundary, a, parameter, mean) %>% 
  pivot_wider(names_from = parameter, values_from = mean)
```

The figures below display the estimates of the $\gamma$ and $\delta$ parameter of the probability weighting function [@prelecProbabilityWeightingFunction1998] fitted to DfE simulated for different parameter values of the generating model.
The estimates are plotted against the switching probability, where each panel represents a distinct boundary type and level, i.e., number of comparisons. 
Grey dots represent agent level estimates, colored dots represent the mean across all agent level estimates.

Most significantly, there is a strong relationship between the switching probability in the generating model and the $\gamma$ parameter.
I.e., large switching probabilities, which are indicative for small sample sequences, lead to larger estimates for $\gamma$. 
The resulting strong curvature leads to a compression of probabilities in the lower and upper range, reflecting underweighting of small probabilities and overweighting of large probabilities. 
This pattern is robust for varying degrees of the boundary level, except for $a = 1$. 
The latter deviation may be explained by the potential scale reduction factors for the respective parameter estimates, which indicate that the MCMC chains did not converge (see [Appendix 3][Appendix 3: Convergence Diagnostics for CPT Parameters]).

@@

# Density Plot

cpt_long %>% 
  filter(parameter == "gamma" | parameter == "delta") %>% 
  ggplot(aes(x = s, y = mean)) +
  geom_density_2d_filled(contour_var = "ndensity") + 
  scale_fill_scico_d(palette = "buda") +
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(limits = c(0,2), breaks = seq(0, 2, .2)) +
  labs(x ="Switching Probability",
       y = "Parameter Value",  
       fill="Density") + 
  facet_grid(parameter~a) + 
  theme_minimal()
```

```{r eval=FALSE, include=FALSE}

# Scatter Plots

# Gamma

cpt_wide %>% 
  ggplot(aes(x = s, y = gamma, color = s)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a) +
  labs(title = expression(paste("Curvature ", gamma)),
       x ="Switching Probability",
       y = expression(gamma), 
       color="Switching Probability") + 
  theme_minimal()

# Delta

cpt_wide %>% 
  ggplot(aes(x = s, y = delta, color = s)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Curvature ", delta)),
       x ="Switching Probability",
       y = expression(delta), 
       color="Switching Probability") + 
  theme_minimal()
```

Below, the resulting probability weighting functions are displayed.

# Weighting Functions

@@

cpt_w <- cpt_wide %>% 
  select(-c(alpha, rho)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>% #
  mutate(w = round(  (delta * ep^gamma)/ ((delta * ep^gamma)+(1-ep)^gamma), 2))

@@

cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x ="Experienced Probability",
       y = "Probability Weighting", 
       color="Switching Probability") + 
  theme_minimal()

cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  facet_wrap(~a) +
  labs(x ="Experienced Probability",
       y = "Probability Weighting", 
       color="Switching Probability") + 
  theme_minimal()
```

The false response rates for different parameter values of the generating model reflect the probability weighting patterns from above. 
That is, the strong curvature resulting from large switching probabilities produces an underweighting of small probabilities. This in turn has the effect that the rarity of an attractive (unattractive) outcome leads to higher rates of choosing the safe (risky) prospect although the risky (safe) prospect had a higher experienced expected value.

# compute false response rates

fr_rates <- choices %>% 
  mutate(ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2), # experienced EV (eEV)
         norm = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) %>% # normative choice according to eEV
  filter(!is.na(norm)) %>% # exclude trials with normative indifferent prospects
  group_by(s, boundary, a, rare, norm, choice) %>% # group correct and incorrect responses
  summarise(n = n()) %>% # absolute numbers 
  mutate(rate = round(n/sum(n), 2), # response rates 
         type = case_when(norm == "A" & choice == "B" ~ "false safe", norm == "B" & choice == "A" ~ "false risky")) %>% filter(!is.na(type))  # remove correct responses

# violin scatter plot

fr_rates %>% 
  ggplot(aes(x = rare, y = rate, color = s)) +
  geom_quasirandom(aes(shape = type), size = 3) +  
  scale_y_continuous(breaks = seq(0, 1, .2)) +
  scale_color_scico(palette = "buda") + 
  scale_shape_manual(values=c(8, 16))+
  labs(x = "Rare Event", 
       y = "False Response Rate", 
       color = "Switching Probability",
       shape = "False Response") + 
  theme_minimal() 
```

```{r eval=FALSE, include=FALSE}
fr_rates %>% 
  ggplot(aes(a, s, fill = rate)) + 
  geom_tile(colour="white", size=0.1) +
  scale_fill_scico(palette = "buda") + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  scale_x_continuous(expand=c(0,0), breaks = seq(1, 5, 1)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(.1, 1, .1)) +
  labs(title = "False Response Rates", 
       x = "a", 
       y= "s", 
       fill = "% False Responses") + 
  theme_minimal()

fr_rates %>% 
  ggplot(aes(s, rate, color = s)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_jitter(size = 3) + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  scale_color_scico(palette = "buda") + 
  labs(title = "False Response Rates", 
       x = "s", 
       y= "% False Responses", 
       color = "a") + 
  theme_minimal()
```

@@

# Density Plot

cpt_wide %>%
  ggplot(aes(x = s, y = alpha)) +
  geom_density_2d_filled(contour_var = "ndensity") + 
  scale_fill_scico_d(palette = "buda") +
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(limits = c(0,2), breaks = seq(0, 2, .2)) +
  labs(x ="Switching Probability",
       y = "Parameter Value",  
       fill="Density") + 
  facet_wrap(~a, nrow = 1) + 
  theme_minimal()
```

```{r}
# value function

## compute values

cpt_v <- cpt_wide %>% 
  select(-c(gamma, delta, rho)) %>% 
  expand_grid(x = seq(0, 20, 2)) %>%  
  mutate(v = round(x^alpha, 2))

cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) + 
  labs(title = "Value Function",
       x ="Objective Value",
       y = "Subjective Value", 
       color="Switching Probability") + 
  theme_minimal()

cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) + 
  facet_wrap(~a) +
  labs(title = "Value Function",
       x ="Objective Value",
       y = "Subjective Value", 
       color="Switching Probability") + 
  theme_minimal()
```

```{r eval=FALSE, include=FALSE}

# alpha
cpt_wide %>% 
  ggplot(aes(x = s, y = alpha, color = s)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(alpha),
       x ="Switching Probability",
       y = expression(alpha), 
       color="Switching Probability") + 
  theme_minimal()

```

@@

# Acknowledgements

@@

We start by noting that the probability of success equals the probability of the subset $B_1 \in \mathcal{G}$ which contains all comparisons $c \in C$ that satisfy the inequation $\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0$.
Since the probability of the subset $B_1$ equals the probability that any of its elements occurs, the probabilities of all $c \in B_1$ are summed up. 
We therefore use the sum operator with the limits $j \in \{1, ..., |B_1|\}$, with $|B_1|$ denoting the cardinality of the set $B_1$, i.e., the number of elements within the set. 
As a major assumption is that the random variables within a sample sequence are iid and independent across sequences, the joint probability of a pair $\overline{X}_{N_X} \cap \overline{Y}_{N_Y}$, i.e., a pair $X_i \cap Y_i$ with a unique combination of the number of times any possible realization of $X$ and $Y$ occurs, is described by the product of their individual probabilities $p(\overline{X}_{N_X})$ and $p(\overline{Y}_{N_Y})$.
Finally, to get these individual probabilities, we must determine the probability that each possible realization indeed occurs the respective number of times in the underlying sample sequence.
This probability is given by the multinomial distribution which reduces to a binomial distribution in the case of only two possible realizations. 
We denote the number of times a realization occurs in a sample sequence with $k_{\omega}$, where $\omega$ signals that the realizations of the random variables are numerical equivalent to the outcomes in the sample space of the prospects.

$$
  \begin{split}
    p(D = 1) 
      & = p \left( \; B_1 = \{c \in C: (\overline{X}_{N_X} - \overline{Y}_{N_Y} > 0)\} \in \mathcal{G} \; \right) \\\\\\ 
      & = \sum_{j=1}^{|B_1|} \left( \;
          p(\overline{X}_{N_Xj})  
          \times 
          p(\overline{Y}_{N_Yj}) \; 
        \right)_{j \, \in \, \{1, ..., |B_1|\}} \\\\\\
      & = \sum_{j=1}^{|B_1|} \left( \; 
        p \left( 
          \frac{1}{N_X} \sum\limits_{i = 1}^{N_X} X(\omega \in \Omega_X)_{ij}
        \right) 
        \times
        p \left( 
          \frac{1}{N_Y} \sum\limits_{i = 1}^{N_Y} Y(\omega \in \Omega_Y)_{ij}
        \right) \; 
        \right)_{j \, \in \, \{1, ..., |B_1|\}} \\\\\\
      & = \sum_{j=1}^{|B_1|} \left( \;
        \left(
          {N_X\choose k_{\omega_{1X}j}, ..., k_{\omega_{nX}j}} \prod_{\omega_{iX} \in \Omega_X} p(\omega_{iX})^{k_{\omega_{iX}j}}
        \right)
        \times
        \left(
        {N_Y\choose k_{\omega_{1Y}j}, ..., k_{\omega_{nY}j} } \prod_{\omega_{iY} \in \Omega_Y} p(\omega_{iY})^{k_{\omega_{iY}j}}
        \right)
        \right)_{j \, \in \, \{1, ..., |B_1|\}} \\\\\\
      & = \sum_{j=1}^{|B_1|} \left( \;
        \left(
          \frac{N_X!}{k_{\omega_{1X}j}! \times ... \times k_{\omega_{nX}j}!} 
          \times 
          p(\omega_{1X})^{k_{\omega_{1X}j}} 
          \times 
          ... 
          \times p(\omega_{nX})^{k_{\omega_{nX}j}}
        \right)
        \times
        \left(
          \frac{N_Y!}{k_{\omega_{1Y}j}! \times ... \times k_{\omega_{nY}j}!} 
          \times 
          p(\omega_{1Y})^{k_{\omega_{1Y}j}} 
          \times 
          ... 
          \times p(\omega_{nY})^{k_{\omega_{nY}j}}
        \right)
      \right)_{j \, \in \, \{1, ..., |B_1|\}}
  \end{split}
$$

@@
